Setting up Data

### 2019-10-08 doc/Main_Report.rmd
removed Peromyscus notes from meta file
added standardized color fill to setup data

---
title: "MLH1 Data Report"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
    toc_float: yes
  pdf_document: default
---

@@

```{r repeatability_rep2, echo=FALSE, include=FALSE, eval=FALSE}
#2. old vs new, (remove duplicates and merge batch0 with old_master_mlh1)
#load, clean up and mat one of the master.MLH1count sheets
#master
old.master.MLH1 = read.csv("~./MLH1repo/data/old_master_MLH1_nov15.csv", header = TRUE)
#remove all with 'X'
#create a real fileName col, add rev.tif,

@@

```{r repeatability_rep3, echo=FALSE, include=FALSE}
chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions.csv", header = TRUE)

#what are the target columns for this excel sheet?
#espically for getting repeat data
#MLH1.dif  same_count  quality

@@

Some papers have documented age effects for male genome-wide recombination rate (cite, one of the Cole papers).

@@

```{r age.hists, echo=FALSE, eval=FALSE,  warning=FALSE}
meta.data$DOB <- as.Date(meta.data$DOB, format= "%m/%d/%Y")#%Y 2000, %y '18,

#make Mouse meta data work
MouseMetaData.wrkin <- meta.data[,c(1,4)]
colnames(MouseMetaData.wrkin) <- c("mouse", "DOB")

MouseMetaData.wrkin <- MouseMetaData.wrkin[!(is.na(MouseMetaData.wrkin$mouse)|MouseMetaData.wrkin$mouse==""),]

#do this for the mouse table instead of the MLH1 table
MLH1.imprv <- merge(MLH1_data, MouseMetaData.wrkin)#don't apply the all parameters

#add euth date for calculating age

MLH1.imprv <- add_euth_date(MLH1.imprv)
#
MLH1.imprv$euth.date2 <- MLH1.imprv$euth.date
MLH1.imprv$euth.date <- as.Date(as.character(MLH1.imprv$euth.date2), format='%d%b%y')
#this is currently broken
MLH1.imprv <- add_age(MLH1.imprv)

@@

pp <- ggplot(MLH1.imprv.male, aes(age.weeks))+geom_histogram()+facet_wrap(~category)+ggtitle("Age Distributions")
pp

#what is this meta.data.file?
#meta.data.file2 <- add_strain(meta.data.file2)
#meta.data.file2 <- add_sex(meta.data.file2)
#meta.data.file2 <- add_category(meta.data.file2)
#meta.data.file2 <- add_euth_date(meta.data.file2)
#meta.data.file2 <- add_age(meta.data.file2)

#qq <- ggplot(meta.data.file2[meta.data.file2$sex == "male",], aes(age.weeks))+geom_histogram()+facet_wrap(~category)+ggtitle("recorded ages of dissected mice")
#qq

@@

```{r age.effect.plot, warning=FALSE, echo=FALSE}

#make sure the right dataset 
male.data <- AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$sex.y == "male") & (AP_mouse_table_w.Ages$category.y != "other"),]
#remove NAs

male.data <- male.data[male.data$Ncells > 10,]

age.effect.plot <- ggplot(male.data, aes(y=mean_co, x=age.weeks, color=strain.y))+geom_point()+ ylim(c(19,34))+
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Age and mean CO per mouse")+facet_wrap(~subsp.y)
age.effect.plot
#todo add line to strain.y

>>>>

---
title: "MLH1 Data Report"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
    toc_float: yes
  pdf_document: default
bibliography: refs_Main_Report.bib
---

@@

```{r repeatability_rep2, echo=FALSE, include=FALSE, eval=FALSE}
#2. old vs new, (remove duplicates and merge batch0 with old_master_mlh1)
#load, clean up and mat one of the master.MLH1count sheets
#master
old.master.MLH1 = read.csv("~./MLH1repo/data/old_master_MLH1_nov15.csv", header = TRUE)
#remove all with 'X'
#create a real fileName col, add rev.tif, 
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")

@@

```{r repeatability_rep3, echo=FALSE, include=FALSE}
chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
#what are the target columns for this excel sheet?
#espically for getting repeat data
#MLH1.dif  same_count  quality

@@

Some papers have documented age effects for male genome-wide recombination rate [@zelazowski2017age].

@@

```{r age.hists, echo=FALSE, eval=FALSE,  warning=FALSE}
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
meta.data$DOB <- as.Date(meta.data$DOB, format= "%m/%d/%Y")#%Y 2000, %y '18,

#make Mouse meta data work .. rename these
MouseMetaData.wrkin <- meta.data[,c(1,4)]
colnames(MouseMetaData.wrkin) <- c("mouse", "DOB")
MouseMetaData.wrkin <- MouseMetaData.wrkin[!(is.na(MouseMetaData.wrkin$mouse)|MouseMetaData.wrkin$mouse==""),]
#do this for the mouse table instead of the MLH1 table
MLH1.imprv <- merge(MLH1_data, MouseMetaData.wrkin)#don't apply the all parameters
#DOB is date
MLH1.imprv <- add_euth_date(MLH1.imprv)
#good euth date..

#make backup date col #add euth date for calculating age
MLH1.imprv$euth.date2 <- MLH1.imprv$euth.date
MLH1.imprv <- add_age(MLH1.imprv)
#age works

@@

pp <- ggplot(MLH1.imprv[MLH1.imprv$sex == "male", ], aes(age.weeks))+geom_histogram()+facet_wrap(~category)+ggtitle("Male Age and MLH1 counts by subspecies")
pp

@@

```{r age.effect.plot, warning=FALSE, echo=FALSE}
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")

AP_mouse_table_w.Ages <- add_category(AP_mouse_table_w.Ages)
#remove NAs
AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[!(is.na(AP_mouse_table_w.Ages$Nmice) | AP_mouse_table_w.Ages$Nmice==""), ]

#make sure the right dataset 
male.data <- AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$sex.y == "male") & (AP_mouse_table_w.Ages$category.y != "other"),]
#lots of NA rows... ugh .. remove them
male.data <- male.data[!(is.na(male.data$Nmice) | male.data$Nmice==""), ]

#other removed, no points,
male.data <- male.data[male.data$Ncells > 10,]

age.effect.plot <- ggplot(male.data, aes(y=mean_co, x=age.weeks, color=(strain.y) ))+geom_point()+ ylim(c(19,34))+
  #trying this
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male Age and mean CO")+facet_wrap(~subsp.y)
age.effect.plot
#no longer has 'other', #yay I added purtty standard colors
#there is an NA
#todo add line to strain.y

### 2019-10-08 src/SetupData.Rmd
added standardized color fill to setup data

```{r mouse.table, echo=FALSE}
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")

@@

#oo isn't including MLH1 means

@@

#merge with DOB
mouse.table_w.Ages <- merge(batch.mouse.count.DF, meta.data, by.y = "mouse", by.x = "Var1", all = TRUE)
AP_mouse_table_w.Ages <- merge(AP_mouse_table, meta.data, by.y = "mouse", by.x = "mouse", all = TRUE)

@@

>>>>

```{r mouse.table, echo=FALSE}
#source("~./MLH1repo/src/CommonFunc_MLH1repo.R")

@@

@@

#subsp in metadata not fully correct
meta.data <- add_strain(meta.data)
meta.data <- add_subsp(meta.data)
meta.data <- add_sex(meta.data)
meta.data <- add_category(meta.data)

#merge with DOB
mouse.table_w.Ages <- merge(batch.mouse.count.DF, meta.data, by.y = "mouse", by.x = "Var1", all = TRUE)
AP_mouse_table_w.Ages <- merge(AP_mouse_table, meta.data, by.y = "mouse", by.x = "mouse", all = TRUE)
#what are the differences between these two

@@

```{r color.scheme}
#color scheme
#mouse colors

colors_of_strains <- c('WSB'= "#56B4E9",'G' ='cadetblue','LEW'= 'lightblue','PERC'= 'blue',
                       #blue

                       'PWD'= 'red','MSM'= 'darkorange', 'MOLF'= 'orange3',#red mucs, orange Molossi
                       'SKIVE'= 'coral1','KAZ'= 'coral4', 'TOM'= 'indianred1','AST' = 'red4',
                       'CZECH'= '#E69F00',

                    'CAST'='green','HMI'= 'forestgreen', #green

                    'SPRET'='gold','SPIC'= 'goldenrod1',#gold?
                    'CAROLI'='gold3',

                    "F1"='grey', "other" = 'grey'
                    )
